---
title: "Stocks with DuckDB"
format: html
---

## Load libraries and data

Load libraries `duckdb`, `tidyverse`, `dbplyr` and `stringr`.

```{r}
#| echo: false
#| warning: false
library(duckdb)
library(tidyverse)
library(stringr)
library(dbplyr)
library(knitr)
```

In this example, we’ll read in some stock price data from CSVs that live in GitHub.

```{r}
#| echo: true

# list of URLs containing our csv files
base_url <- "https://raw.githubusercontent.com/Robot-Wealth/r-quant-recipes/master/data/"
stocks <- c("AAPL", "BA", "CAT", "IBM", "MSFT")
urls <- str_glue("{base_url}{stocks}.csv")
```

Read data into a dataframe and display.

```{r}
# read data into a dataframe
prices_df <- urls %>%
  map_dfr(read_csv, col_types = "Dddddddc") %>%
  arrange(Date, Ticker)

head(prices_df) |>
  kable()
```


```{r}
# to make a persistent db on disk (and to connect to it again later) do:
con <- dbConnect(duckdb::duckdb(dbdir = "mydatabase.duckdb"))

# write our prices data to duckdb table
table_name <- "prices"
duckdb::dbWriteTable(con, table_name, prices_df, overwrite = TRUE)

# remove prices_df
rm(prices_df)

dbDisconnect(con)

```

### 3. Query some data in our db
```{r}
con <- dbConnect(duckdb::duckdb(dbdir = "mydatabase.duckdb"))
tbl(con, "prices") %>%
  group_by(Ticker) %>%
  summarise(
    Count = n(),
    From = first(Date),
    To = last(Date)
  ) |>
  kable()
```

You can also send SQL to be executed on the database. If you want to get the result back in your R session, use dbGetQuery, otherwise use dbSendQuery:
```{r}
result <- dbGetQuery(con, paste0("SELECT * FROM prices LIMIT 5"))
result |>
  kable()
```
You can do the operation above by writing an equivalent dplyr operation and piping it into collect:

```{r}
tbl(con, "prices") %>%
  head(5) |>
  collect() |>
  kable()
dbDisconnect(con)
```
```{r}
# Basic TWAP approximation

con <- dbConnect(duckdb::duckdb(dbdir = "mydatabase.duckdb"))

# will work
tbl(con, "prices") |>
   dplyr::mutate(TWAP = (Open + High + Low + Close)/4) |>
   head() |>
  kable()

dbDisconnect(con)
```

```{r}
con <- dbConnect(duckdb::duckdb(dbdir = "mydatabase.duckdb"))
# 60-day moving average
affected_rows <- dbExecute(con,
    'ALTER TABLE prices ADD COLUMN IF NOT EXISTS MA60 DOUBLE;
     UPDATE prices SET MA60 = t2.MA60
     FROM (
        SELECT Ticker, Date, AVG("Adj Close") OVER (
            PARTITION BY Ticker ORDER BY Date ROWS BETWEEN 59 PRECEDING AND CURRENT ROW
        ) AS MA60
    FROM prices
    ) t2
WHERE prices.Ticker = t2.Ticker AND prices.Date = t2.Date;'
)

# prices <- tbl(con, "prices")

# prices |>
# group_by(Ticker) |>
# window_order(Date) |>
# window_frame(-59, 0) |>
# mutate(MA60 = mean(`Adj Close`)) |>
# ungroup()

affected_rows

tbl(con, "prices") |> head() |> kable()

dbDisconnect(con)
```
```{r}
con <- dbConnect(duckdb::duckdb(dbdir = "mydatabase.duckdb"))

# Set chart options
options(repr.plot.width = 14, repr.plot.height=7)
theme_set(theme_bw())
theme_update(text = element_text(size = 12))

# plot
tbl(con, "prices") %>%
  select(Date, Ticker, `Adj Close`, MA60) %>%
  pivot_longer(cols = c(`Adj Close`, MA60), names_to = "Type", values_to = "Price") %>%
  ggplot(aes(x = Date, y = Price, colour = Type)) +
    geom_line() +
    facet_wrap(~Ticker) +
    labs(title = "Closing prices and their 60-day moving average")

dbDisconnect(con)
```

